Open-Set Domain Adaptation through Self-Supervision

### 2022-02-13 paper/main.tex
center loss statement and minor fixes

The application of the rotation recognition pre-text task allows to effectively favor and force the model to learn
domain-independent patterns, which are crucial to perform the novelty detection in a cross-domain fashion and, moreover,
to successfully perform the domain alignment.
To provide an explanation of why this applies, we can think that a rotation classifier needs to focus on discriminative patterns to
successfully perform the rotation predictions, such as shapes, edges, and high-level object relative position like the position of the eyes
w.r.t nose.
The method and its effectiveness is further illustrated in \cite{OldROS},
and further improvements will be discussed in \ref{sec:future_work}.

@@

Here ${\bf v_i}$ is the output of the penultimate layer of $R_1$,
$\gamma({\bf z}_i)$ is the centroid of the features associated to class $i$
(notice that the centroid is relative to a different rotation class $i$ in the multi-head variant),
$||\cdot||_2^2$ is the $l$-2 norm and $\lambda$ is the weight associated with the center loss term.

@@

As we have already seen in section \ref{sec:method}, the network is composed of a feature extractor $E$, and two heads, $C_{1/2}$ and $R_{1/2}$.
For the feature extraction, we employ a ResNet-18\cite{ResNet18} pre-trained on ImageNet.
All the experiments are run training with a batch size equals to 32 and using a stochastic gradient descent optimization strategy
with a weight decay of 0.0005 and a momentum of 0.9.
The learning rate is set to 0.001 in the cases where we use a single-head rotation classifier as $R_1$, and 0.003 in the multi-head case.
The reason of this choice lies on the fact that, with a multi-head rotation classifier, we have to train 45 heads instead of a single one,
so the training process needs to be speeded up.
Moreover, we set a learning rate 10 times lower for the weights of the feature extractor and $C_{1/2}$.
We also decrease the learning rate using a step learning rate scheduler, which reduces it by a factor of 10 after 90\% of total epochs.
All models are trained for 50 epochs for the step one and for 25 epochs for the step two.
For models using the center loss a centroid learning rate of 0.001 is used.
The other hyperparameters are analyzed performing an ablation on them,
details can be found on the section \ref{sec:ablation_study} and the results of the ablation are
reported on appendix \ref{sec:tables}.

@@

{\it Classifier $R_1$, $R_2$}: 
they are seen as discriminators composed by a linear layer $1024\to 256$,
followed by a batch normalization layer and a leaky relu activation.
The last layer is then a linear one for the actual classification $256\to 4$.
In the case of multi-head rotation classifier,
the discriminator is then composed of $|\mathcal{C}_s|$ different linear layers, both for the first and the last one.
Moreover, the 256-dimensional output are the features used by the center loss for inferring class centroids.
$R_2$ is always a single-head classifier, even if using a multi-head rotation classifier.

>>>>

The application of the rotation recognition pre-text task allows to effectively favor and force the model to learn
domain-independent patterns, which are crucial to perform the novelty detection in a cross-domain fashion and, moreover,
to successfully perform the domain alignment.
To provide an explanation of why this applies, we can think that a rotation classifier needs to focus on discriminative patterns to
successfully perform the rotation predictions, such as shapes, edges, and high-level object relative position like the position of the eyes
w.r.t nose.
The method and its effectiveness is further illustrated in \cite{OldROS},
and further possible improvements will be discussed in section \ref{sec:future_work}.

@@

Here ${\bf v_i}$ is the output of the penultimate layer of $R_1$,
$\gamma({\bf z}_i)$ is the centroid of the features associated to class $i$
(notice that the centroid is relative to a different rotation class $i$ in the multi-head variant),
$||\cdot||_2^2$ is the $l$-2 norm and $\lambda$ is the weight associated with the center loss term.
The additional center loss contribution should favor the learning of hidden features which have
a smaller intra-class variance, resulting in a more discriminative representation
when combined with the standard cross-entropy loss,
which in contrast tries to enlarge the distance between the means of the classes.

@@

As we have already seen in section \ref{sec:method}, the network is composed of a feature extractor $E$, and two heads, $C_{1/2}$ and $R_{1/2}$.
For the feature extraction, we employ a ResNet-18\cite{ResNet18} pre-trained on ImageNet.
All the experiments are run training with a batch size equals to 32 and using a stochastic gradient descent optimization strategy
with a weight decay of 0.0005 and a momentum of 0.9.
The learning rate is set to 0.001 in the cases where we use a single-head rotation classifier as $R_1$, and 0.003 in the multi-head case.
The reason of this choice lies on the fact that, with a multi-head rotation classifier, we have to train 45 heads instead of a single one,
so the training process needs to be speeded up.
Moreover, we set a learning rate 10 times lower for the weights of the feature extractor and $C_{1/2}$.
We also decrease the learning rate using a step learning rate scheduler, which reduces it by a factor of 10 after 90\% of total epochs.
All models are trained for 50 epochs for the step one and for 25 epochs for the step two.
For models using the center loss, a centroid learning rate of 0.001 is used.
The other hyperparameters are analyzed performing an ablation on them,
details can be found on the section \ref{sec:ablation_study} and the results of the ablation are
reported on appendix \ref{sec:tables}.

@@

{\it Classifier $R_1$, $R_2$}: 
they are seen as discriminators composed by a linear layer ($1024\to 256$),
followed by a batch normalization layer and a leaky relu activation.
The last layer is then a linear one for the actual classification ($256\to 4$).
In the case of multi-head rotation classifier,
the discriminator is then composed of $|\mathcal{C}_s|$ different linear layers, both for the first and the last one.
Moreover, the 256-dimensional output are the features used by the center loss for inferring class centroids.
$R_2$ is always a single-head discriminator, even if using the multi-head variant.
